Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2016-03-02 draft.tex
Removed statement about work in progress from cover page
change time subscripts to superscripts through defn 1

\date{\today}
\begin{document}
\begin{titlepage}
\maketitle
\thispagestyle{empty}
\center{This article is a working paper and is subject to future revision during the peer-review process. The authors welcome any comments, and may be reached by e-mail at benjamin.a.graham@usc.edu.}

@@

Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates. This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} U_{it} \de^{t-1}$.

@@

\item $p$ chooses a level of resource expenditure $R_{pt}\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.

\item $c$ chooses a level of resource expenditure $R_{ct}\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.

@@

\noindent Stage game payoffs for players $s$ and $g$ in period $n$, gross of investments by players $p$ and $c$, are:
\vspace{2pt}
\begin{center}
\begin{tabular}{|  c| c | c | c |} 
\hline
  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
	Fight & $\Omega_t$ &$\Omega_t$&$W_{gt}, L_{st}$ \\ \hline
	Status Quo& $\Omega_t$& $Q_{gt}, Q_{st}$ &$W_{gt}, L_{st}$ \\ \hline
	Cede&$L_{gt}, W_{st}$& $L_{gt}, W_{st}$ &$Q_{gt}, Q_{st}$ \\ \hline
 \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
\end{tabular}
\end{center}
\vspace{2.5pt}

@@

For probabilities $p_1$ of outright victory and $1-p_1$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_{it} \equiv (p_1 (W_{it}-\zeta_i), \left(1-p_1\right)( L_{it}-\zeta_i))$. $\Omega_t \equiv (\omega_{gt}, \omega_{st}).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$. Note that the structure of the war lottery along with the assumption that the outside actors influence the payoffs only and not the probabilities of winning imply that any way in which the patron influences the payoffs from war can be counteracted by the international community through its own investments.} Players are assumed to approach war lotteries as expected values.

If both $s$ and $g$ play Status Quo, then the status quo persists. The status quo payoffs for the secessionists are modeled as a state variable. The transition of this state variable from one period to the next involves an increase by the amount of the patron's investment and an automatic reduction by $\mu$.\footnote{Note that it is also possible for the international community to make an investment in the secessionists' status quo payoffs, but we will show that they have no incentive to do so.} The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction. As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists. Therefore, the payoffs from the status quo, in the absence of other actions, are given by $q_{s,t+1} = Q_{s,t} - \mu$. Lower case letters denote beginning-of-period quantities net of investments by outside actors, while upper case letters denote end-of-period quantities gross of such investments.

Other payoffs are also modeled as state variables. However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, e.g. $L_{s,t} = l_{s,t} + R_{p,t} + R_{c,t}$ and $l_{s,t+1} = L_{s,t}$, likewise for player $g$ and for the payoffs from winning for both players and the status quo for player $g$.

@@

At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game. For example, if $p$ expends resources $R_{pt}$ to increase the secessionists' status quo payoffs, the beginning-of-period payoff $q_{st}$ becomes end-of-period payoff $Q_{st} = q_{st} + R_{pt}$. We can write analogous expressions for $q_{gt}, \ l_{st},\ l_{gt},\ w_{st}$, and $w_{st}$, and the international community can also make investments to increase whichever of these payoffs they like. $B_{pt}$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.

@@

Player $p$'s payoffs in period $t$ are $U_{pt}= -\alpha X+\lambda Y -R_{pt}$, while the payoffs for player $c$ are $U_{ct}= \beta X-\nu Y -R_{ct}$, with both payoffs denoted in currency units. Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.

@@

\begin{enumerate}
\item \textit{For both players $g$ and $s$, $Q_{i0} \geq L_{i0}$: remaining in the status quo is better than ceding at the beginning of the game.}

\item \textit{For both players $g$ and $s$, $\frac {Q_{i0}}{1-\delta} \geq  -\zeta_i+\frac{W_{i0}(p_{1i}) + L_{i0}(1-p_{1i})}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}

@@

\item  \textit{$B_{p1} \geq\frac{\beta}{1-\delta} - \left(Q_{s0} -\mu - L_{s0} \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}

\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6). If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be the binding constraint.}

@@

The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors. That is, $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.\footnote{In the case of $w_{st}$ and $w_{gt}$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} Strategies for the patron and the international community are how much to invest in each of the six state variables. However, positive investments in some of the state variables can be ruled out by preference assumptions.

@@

\begin{enumerate}
	\item The patron invests $R_{pt} = \max\left\{\frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right),0\right\}$ to augment $q_{st}$. Player $c$ invests $R_{ct} =0$ in $l_{st}$.

If $R_{pt} < \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ player $c$ invests $R_{ct} = l_{st} - \left(q_{st} + R_{pt} \right) + \ve$, for $\ve$ small, in $l_{st}$. Cede becomes the unique best response for the secessionists.

	\item As long as $R_{pt} \leq \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$, player $c$ invests $R_{ct} = R_{pt} - \left( Q_{g,t-1} - L_{g,t-1}\right)$ in $q_{gt}$. Note that the outcome is that investments in $l_{gt}$ and $q_{gt}$ are zero.

If $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$, player $c$ invests $R_{ct} = 0$. Cede becomes the unique best response for the government.

	\item As long as $R_{pt} \leq \frac{1}{p_{1s}} \left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right]$ to augment $w_{st}$, player $c$ invests $R_{ct} = R_{pt}p_{1s} + \mu - Q_{s,t-1} + \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ in $q_{st}$. Similar to case two, the outcome is that investments in $w_{st}$ and $q_{st}$ are zero.

If $R_{pt} > \frac{1}{p_{1s}}\left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right] $ to augment $w_{st}$, player $c$ will choose $R_{ct}=0$. Fight becomes the unique best response for the secessionists.
\end{enumerate}
\end{definition}

@@

\begin{lemma}
$Q_{it} \geq L_{it}$ and $\frac {Q_{it}}{1-\delta} \geq  -\zeta_i+\frac{W_{it}(p_{1i}) + L_{it}(1-p_{1i})}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for (Status Quo,Status Quo) to be the outcome of Stage 3 in period $t \ \forall t$. 
\end{lemma}

@@

Recall that either outside actor could invest toward increasing any of the six state variables in a period: $q_{st},q_{gt},w_{st},w_{gt},l_{st},$ and $l_{gt}$. However, each outside actor only considers investments in three of the state variables as treated in the three cases in Definition 2. The remainder of the proof addresses each set of matched investment decisions in turn.

@@

Consider period 1. When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest. That is, in equilibrium, $R_{p1} = \frac{\beta}{1-\de} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment. Thus we need Restriction (5): $B_{p1} \geq\frac{\beta}{1-\de} - \left(Q_{s0} - \mu - L_{s0} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.

In periods $t > 1$, there are two cases to consider. Either (a) the buffer created in period $t-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$. In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$. Restrictions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$. However, because in each period $q_{st}$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a). Again, assuming Restriction (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction 6).

@@

In Section~\ref{sec:mainproof}, we considered the outside actors' abilities to make investments to increase the various payoffs of the home state government and the secessionists. Player $c$, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e. sanctions are equivalent to $R_{ct} < 0$). Note that this may be particularly effective if $c$ is a large coalition of states acting in concert.

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo (i.e. $q_{s,t+1} = Q_{s,t} - \mu -\sigma$ where $\sigma$ is the reduction due to sanctions).  In order for sanctions to lead to ceding by the secessionists, the following are required:

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e. $q_{s,t+1} = Q_{s,t} - \mu -\sigma$ \emph{and} $p_{1s}\left(\sigma\right)$ $\left(p_{1g}\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$). The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
\end{proposition}

@@

{\bf Proof of Lemma 1} \\
Recall that the upper case notation represents payoffs gross of investments by the outside players. Without loss of generality, consider the incentives to deviate for player $s$ in period $t$ given that the other three players play their equilibrium actions. Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{st} \ \forall t$, so the continuation value is $\frac{Q_{st}}{1-\delta}$. Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{st}$ in each period, is $\frac{L_{st}}{1-\delta}$. The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, and $L_s$ forever with probability $1-p_1$. Thus the continuation value is $-\zeta_s +\frac{W_{st}(p_1)+ L_{st}(1-p_1)}{1-\delta}$.

Given that the continuation value from Status Quo is higher than both that from playing Cede and Fight, player $s$ will play Status Quo. The argument for player $g$ is symmetric. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 2} \\
In the status quo, we have $U_{pt}^{SQ}= -R_p^{SQ}$ and $U_{ct}^{SQ}= -R_c^{SQ}$. If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pt}^{RU}= -\alpha -R_p^{RU}$ and $U_{ct}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.

@@

Player $c$ would be able to induce the secessionists to play Cede---e.g. counter an addition to $q_{st}$ with an investment to $l_{st}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} > Q_{s,t-1} -\mu + R_{pt} - L_{s,t-1}$.\footnote{Here we assume that the payoff relationship at $t$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} For this to work, the investment required of player $c$ must be no more than the gain player $c$ receives from investing.

On the other hand, if $\frac{\beta + R_c^{SQ}}{1 -\de} \leq Q_{s,t-1} -\mu + R_{pt} - L_{s,t-1}$, the patron's investment creates a large enough difference in $Q_{st} - L_{st}$ (net of player $c$'s investment) so that player $c$ will not find contesting the status quo to be in its interest. Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$.

Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} > L_{s,t-1}$. Combining this with $R_c^{SQ} =0$, a tighter condition for preventing player $c$ from inducing reunification is $\frac{\beta}{1 -\de} + \mu \leq R_{pt}$.

Since the patron is willing to invest up to $\frac{\alpha}{1 -\de}$, $\frac{\beta}{1 -\de} + \mu \leq \frac{\alpha}{1 -\de}$ is a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 3} \\
Again, in the status quo, we have $U_{pt}^{SQ}= -R_p^{SQ}$ and $U_{ct}^{SQ}= -R_c^{SQ}$. If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pt}^{REC}= \lambda -R_p^{REC}$ and $U_{ct}^{REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.

@@

If $\frac{\nu}{1-\de} < L_{g,t-1} + R_{pt} - Q_{g,t-1}$, the patron's investment in $l_{gt}$ creates a large enough difference in $L_{st} - Q_{st}$ (net of player $c$'s investment) so that player $c$ will not find it worthwhile to re-establish the status quo outcome.

On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g. counter an addition to $l_{gt}$ with an investment to $q_{gt}$---as long as $\frac{\nu}{1-\de} \geq L_{g,t-1} + R_{pt} - Q_{g,t-1}$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.}

Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} > L_{s,t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_{pt}$.

@@

If $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  < \omega_{st}\left(R_{pt}\right) - Q_{st}$, the patron's investment in $\omega_{st}$ creates a large enough difference between the payoffs from fighting and the status quo so that player $c$ will not find it worthwhile to re-establish the status quo outcome.

On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce fighting---e.g. counter an addition to $\omega_{st}$ with an investment to $q_{st}$---as long as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq \omega_{st}\left(R_{pt}\right) - Q_{st}$.

>>>>

\date{\today}
\begin{document}
\begin{titlepage}
\maketitle
\thispagestyle{empty}
%\center{This article is a working paper and is subject to future revision during the peer-review process. The authors welcome any comments, and may be reached by e-mail at benjamin.a.graham@usc.edu.}

@@

Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates. This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} U_i^t \de^{t-1}$.

@@

\item $p$ chooses a level of resource expenditure $R_p^t\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.

\item $c$ chooses a level of resource expenditure $R_c^t\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.

@@

\noindent Stage game payoffs for players $s$ and $g$ in period $n$, gross of investments by players $p$ and $c$, are:
\vspace{2pt}
\begin{center}
\begin{tabular}{|  c| c | c | c |} 
\hline
  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
	Fight & $\Omega^t$ &$\Omega^t$&$W_g^t, L_s^t$ \\ \hline
	Status Quo& $\Omega^t$& $Q_g^t, Q_s^t$ &$W_g^t, L_s^t$ \\ \hline
	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$Q_g^t, Q_s^t$ \\ \hline
 \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
\end{tabular}
\end{center}
\vspace{2.5pt}

@@

For probabilities $p_1$ of outright victory and $1-p_1$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_i^t \equiv (p_1 (W_i^t-\zeta_i), \left(1-p_1\right)( L_i^t-\zeta_i))$. $\Omega^t \equiv (\omega_g^t, \omega_s^t).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$. Note that the structure of the war lottery along with the assumption that the outside actors influence the payoffs only and not the probabilities of winning imply that any way in which the patron influences the payoffs from war can be counteracted by the international community through its own investments.} Players are assumed to approach war lotteries as expected values.

If both $s$ and $g$ play Status Quo, then the status quo persists. The status quo payoffs for the secessionists are modeled as a state variable. The transition of this state variable from one period to the next involves an increase by the amount of the patron's investment and an automatic reduction by $\mu$.\footnote{Note that it is also possible for the international community to make an investment in the secessionists' status quo payoffs, but we will show that they have no incentive to do so.} The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction. As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists. Therefore, the payoffs from the status quo, in the absence of other actions, are given by $q_s^{t+1} = Q_s^t - \mu$. Lower case letters denote beginning-of-period quantities net of investments by outside actors, while upper case letters denote end-of-period quantities gross of such investments.

Other payoffs are also modeled as state variables. However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, e.g. $L_s^t = l_s^t + R_p^t + R_c^t$ and $l_s^{t+1} = L_s^t$, likewise for player $g$ and for the payoffs from winning for both players and the status quo for player $g$.

@@

At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game. For example, if $p$ expends resources $R_p^t$ to increase the secessionists' status quo payoffs, the beginning-of-period payoff $q_s^t$ becomes end-of-period payoff $Q_s^t = q_s^t + R_p^t$. We can write analogous expressions for $q_g^t, \ l_s^t,\ l_g^t,\ w_s^t$, and $w_g^t$, and the international community can also make investments to increase whichever of these payoffs they like. $B_p^t$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.

@@

Player $p$'s payoffs in period $t$ are $U_p^t= -\alpha X+\lambda Y -R_p^t$, while the payoffs for player $c$ are $U_c^t= \beta X-\nu Y -R_c^t$, with both payoffs denoted in currency units. Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.

@@

\begin{enumerate}
\item \textit{For both players $g$ and $s$, $Q_i^0 \geq L_i^0$: remaining in the status quo is better than ceding at the beginning of the game.}

\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq  -\zeta_i+\frac{W_i^0(p_{1i}) + L_i^0(1-p_{1i})}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}

@@

\item  \textit{$B_p^1 \geq\frac{\beta}{1-\delta} - \left(Q_s^0 -\mu - L_s^0 \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}

\item \textit{$B_p^t \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6). If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be the binding constraint.}

@@

The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors. That is, $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.\footnote{In the case of $w_s^t$ and $w_g^t$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} Strategies for the patron and the international community are how much to invest in each of the six state variables. However, positive investments in some of the state variables can be ruled out by preference assumptions.

@@

\begin{enumerate}
	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right),0\right\}$ to augment $q_s^t$. Player $c$ invests $R_c^t =0$ in $l_s^t$.

If $R_p^t < \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ player $c$ invests $R_c^t = l_s^t - \left(q_s^t + R_p^t \right) + \ve$, for $\ve$ small, in $l_s^t$. Cede becomes the unique best response for the secessionists.

	\item As long as $R_p^t \leq \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = R_p^t - \left( Q_{g,t-1} - L_{g,t-1}\right)$ in $q_g^t$. Note that the outcome is that investments in $l_g^t$ and $q_g^t$ are zero.

If $R_p^t > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = 0$. Cede becomes the unique best response for the government.

	\item As long as $R_p^t \leq \frac{1}{p_{1s}} \left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right]$ to augment $w_s^t$, player $c$ invests $R_c^t = R_p^tp_{1s} + \mu - Q_{s,t-1} + \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ in $q_s^t$. Similar to case two, the outcome is that investments in $w_s^t$ and $q_s^t$ are zero.

If $R_p^t > \frac{1}{p_{1s}}\left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right] $ to augment $w_s^t$, player $c$ will choose $R_c^t=0$. Fight becomes the unique best response for the secessionists.
\end{enumerate}
\end{definition}

@@

\begin{lemma}
$Q_i^t \geq L_i^t$ and $\frac {Q_i^t}{1-\delta} \geq  -\zeta_i+\frac{W_i^t(p_{1i}) + L_i^t(1-p_{1i})}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for (Status Quo,Status Quo) to be the outcome of Stage 3 in period $t \ \forall t$. 
\end{lemma}

@@

Recall that either outside actor could invest toward increasing any of the six state variables in a period: $q_s^t,q_g^t,w_s^t,w_g^t,l_s^t,$ and $l_g^t$. However, each outside actor only considers investments in three of the state variables as treated in the three cases in Definition 2. The remainder of the proof addresses each set of matched investment decisions in turn.

@@

Consider period 1. When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest. That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment. Thus we need Restriction (5): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.

In periods $t > 1$, there are two cases to consider. Either (a) the buffer created in period $t-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$. In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$. Restrictions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$. However, because in each period $q_s^t$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a). Again, assuming Restriction (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction 6).

@@

In Section~\ref{sec:mainproof}, we considered the outside actors' abilities to make investments to increase the various payoffs of the home state government and the secessionists. Player $c$, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e. sanctions are equivalent to $R_c^t < 0$). Note that this may be particularly effective if $c$ is a large coalition of states acting in concert.

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo (i.e. $q_s^{t+1} = Q_s^t - \mu -\sigma$ where $\sigma$ is the reduction due to sanctions).  In order for sanctions to lead to ceding by the secessionists, the following are required:

@@

\begin{proposition}
Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e. $q_s^{t+1} = Q_s^t - \mu -\sigma$ \emph{and} $p_{1s}\left(\sigma\right)$ $\left(p_{1g}\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$). The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
\end{proposition}

@@

{\bf Proof of Lemma 1} \\
Recall that the upper case notation represents payoffs gross of investments by the outside players. Without loss of generality, consider the incentives to deviate for player $s$ in period $t$ given that the other three players play their equilibrium actions. Given the stationary equilibrium actions, playing Status Quo will lead to $Q_s^t \ \forall t$, so the continuation value is $\frac{Q_s^t}{1-\delta}$. Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_s^t$ in each period, is $\frac{L_s^t}{1-\delta}$. The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, and $L_s$ forever with probability $1-p_1$. Thus the continuation value is $-\zeta_s +\frac{W_s^t(p_1)+ L_s^t(1-p_1)}{1-\delta}$.

Given that the continuation value from Status Quo is higher than both that from playing Cede and Fight, player $s$ will play Status Quo. The argument for player $g$ is symmetric. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 2} \\
In the status quo, we have $U_p^t^{SQ}= -R_p^{SQ}$ and $U_c^t^{SQ}= -R_c^{SQ}$. If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_p^t^{RU}= -\alpha -R_p^{RU}$ and $U_c^t^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.

@@

Player $c$ would be able to induce the secessionists to play Cede---e.g. counter an addition to $q_s^t$ with an investment to $l_s^t$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} > Q_{s,t-1} -\mu + R_p^t - L_{s,t-1}$.\footnote{Here we assume that the payoff relationship at $t$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} For this to work, the investment required of player $c$ must be no more than the gain player $c$ receives from investing.

On the other hand, if $\frac{\beta + R_c^{SQ}}{1 -\de} \leq Q_{s,t-1} -\mu + R_p^t - L_{s,t-1}$, the patron's investment creates a large enough difference in $Q_s^t - L_s^t$ (net of player $c$'s investment) so that player $c$ will not find contesting the status quo to be in its interest. Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$.

Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} > L_{s,t-1}$. Combining this with $R_c^{SQ} =0$, a tighter condition for preventing player $c$ from inducing reunification is $\frac{\beta}{1 -\de} + \mu \leq R_p^t$.

Since the patron is willing to invest up to $\frac{\alpha}{1 -\de}$, $\frac{\beta}{1 -\de} + \mu \leq \frac{\alpha}{1 -\de}$ is a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 3} \\
Again, in the status quo, we have $U_p^t^{SQ}= -R_p^{SQ}$ and $U_c^t^{SQ}= -R_c^{SQ}$. If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_p^t^{REC}= \lambda -R_p^{REC}$ and $U_c^t^{REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.

@@

If $\frac{\nu}{1-\de} < L_{g,t-1} + R_p^t - Q_{g,t-1}$, the patron's investment in $l_g^t$ creates a large enough difference in $L_s^t - Q_s^t$ (net of player $c$'s investment) so that player $c$ will not find it worthwhile to re-establish the status quo outcome.

On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g. counter an addition to $l_g^t$ with an investment to $q_g^t$---as long as $\frac{\nu}{1-\de} \geq L_{g,t-1} + R_p^t - Q_{g,t-1}$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.}

Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} > L_{s,t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_p^t$.

@@

If $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  < \omega_s^t\left(R_p^t\right) - Q_s^t$, the patron's investment in $\omega_s^t$ creates a large enough difference between the payoffs from fighting and the status quo so that player $c$ will not find it worthwhile to re-establish the status quo outcome.

On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce fighting---e.g. counter an addition to $\omega_s^t$ with an investment to $q_s^t$---as long as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq \omega_s^t\left(R_p^t\right) - Q_s^t$.
